### 2013-04-22 2/literature_review.tex
added 201304_Apr22

\section{The nature of Polarimetric SAR: EM Wave Polarization}

Remote sensing introduction paragraph (section 2.1.1)

SAR introduction paragraph (section 2.1.2)

EM Wave polarization (Maxwell law leads to wave equations section 2.1.3.1, choice of origin which leads to polarization basis change eq 2.22)

SAR Polarimetry: and SAR as a special case of POLSAR (send and receive polarizations: 2.1.3.5)

Full Polarimetric SAR and target polarimetric response (also 2.1.3.5)

Partial Polarimetric SAR and EM wave polarization (2.1.3.4)

SAR stochastics nature (new)

\subsection{Representations of EM Wave Polarizations (Linear, Circular ...)}

\subsubsection{Plane wave and solutions to the EM Wave equations}

The general solution to the PDE EM wave equation is a linear superposition of waves each having the form
\begin{equation}
\vec{E}(\vec{r},t)=g(\phi(\vec{r},t)) = g(\omega t - \vec{k} \cdot \vec{r})
\end{equation} 
and
\begin{equation}
\vec{B}(\vec{r},t)=g(\phi(\vec{r},t)) = g(\omega t - \vec{k} \cdot \vec{r})
\end{equation}
where
	$g$ is any well-behaved function of a dimensionless argument $\phi$,
	%for virtually any well-behaved function g of dimensionless argument φ, where
	$\omega$ is the angular frequency (in radians per second), and
	$\vec{k}= (k_x,k_y,k_z)$ is the wave vector (in radians per meter).

%Although the function $g$ can be, and often is, a monochromatic sine wave, it does not have to be. 
%In practice, $g$ cannot have infinite periodicity because any real electromagnetic wave must always have a finite extent in time and space. 
%Thus based on the theory of Fourier decomposition, a real wave must consist of the superposition of an infinite set of sinusoidal frequencies.
In addition to the relations above, the wave vector and the angular frequency are not independent; they must adhere to the dispersion relation:
\begin{equation}
k = |\vec{k}| = \frac{\omega}{c} = \frac{2\pi}{\lambda}
\end{equation}
where 
	$k$ is the wave number and $\lambda$ is the wavelength.

Since any real wave can be considered as a superpositions of several sinusoidal waves, the simplest solution to consider for the EM Wave equations is:
\begin{equation}
\vec{E}(\vec{r},t)= \Re \left( \vec{E}(\vec{r}) e^{i \omega t} \right)
\end{equation}
where
	$\omega = 2 \pi f$ is the angular frequency,
	$f$ is the wave frequency.
Assuming the most common case of EM waves travelling in a planar fashion, which is a good approximation of most light waves. 
For plane waves Maxwell's equations, specifically Gauss's laws, imposes a transversality requirement that the electric and magnetic field be perpendicular to the direction of propagation and to each other. 
Conventionally, when considering polarization, the electric field vector  is described and the magnetic field is ignored since it is perpendicular  to the electric field and proportional to it. 
%Since the magnetic field is dependent on the electric field, 
Thus it is sufficient to just consider electric field, which equals:
\begin{equation}
\vec{E}(\vec{r}) = \vec{E^0} e^{-i \vec{k} \cdot \vec{r}} %= \vec{E^0} e^{-ikz}
\end{equation}

Combining the two equations we have:
\begin{equation}
\vec{E}(\vec{r},t) 
	= \vec{E^0} \Re \left( e^{i (\omega t - \vec{k} \cdot \vec{r})} \right)
	= \vec{E^0} \cos(wt - \vec{k} \cdot \vec{r}) %\vec{E^0} \cos{ \left( wt - kz + \phi_0 \right)}
\end{equation}

If we define the $z$ direction as the travelling direction of the EM wave,
%	and x direction is the direction of $\vec{E}$,
%	then by Faraday's Law, the magentic field $\vec{B}$ lies in y direction, 
%		and is related to the electric vector by $c^2 \frac{\partial B}{\partial z} = \frac{\partial E}{\partial t}$
	and the electric field is fluctuating in the $(x,y)$ plane,
	because the divergence of the electric and magnetic fields are zero, 
		there will be no fields in the direction of propagation.
Project this into $x$ and $y$ directions:
\begin{equation}
\vec{E}(z,t) 
= 
\left(
	\begin{array} {c}
		E_x^0 \cos{ \left( wt - kz  + \phi_x^0 \right) } \\
		E_y^0 \cos{ \left( wt - kz  + \phi_y^0 \right)} \\
		0
	\end{array}
\right)
= 
\left(
	\begin{array} {c}
		\xi_x(t) \\
		\xi_y(t) \\
		0
	\end{array}
\right)
\end{equation}
where $\phi_0$ is the initial phase angle, that is dependent on the choice of space-time origin.

\subsubsection{Linear, Circular and Elliptical Polarization}

The shape traced out in a fixed projective plane by the electric vector as the plane wave passes over it is a description of the polarization state. 
This shape is characterized by the formula
\begin{equation}
 \frac{\xi_x^2(t)}{|E_x^0|^2} + \frac{\xi_y^2(t)}{|E_y^0|^2} - 2 \frac{\xi_x \xi_y}{E_x^0 E_y^0} \cos (\Delta \phi) = \sin^2(\Delta \phi)
\end{equation}
where
	$\Delta \phi = \phi_y^0 - \phi_x^0$
The formula characterizes what is termed as the polarization ellipse.

%In the leftmost figure above, 
When the two orthogonal components are in phase, i.e. $\phi_x^0 = \phi_y^0$, the ratio of the strengths of the two components is constant.
Thus the direction of the electric vector (the vector sum of these two components) is constant. 
Since the tip of the vector traces out a single line in the plane, this special case is called linear polarization. 
The direction of this line depends on the relative amplitudes of the two components.
%In the middle figure,

When the two orthogonal components have exactly the same amplitude and are exactly ninety degrees out of phase, i.e. $E_x^0=E_y^0$ and $ | \phi_x^0 - \phi_y^0 | = \pi/2$. 
%There are two possible phase relationships that satisfy this requirement: the x component can be ninety degrees ahead of the y  component or it can be ninety degrees behind the y component. 
In this special case the electric vector traces out a circle in the plane, so this special case is called circular polarization. 
The direction the field's rotation depends on the signed of the difference between the two phases. 
These cases are called right-hand circular polarization and left-hand circular polarization, 
	depending on which way the electric vector rotates and the chosen convention.

The general case is when the two components are not in phase 
	and either do not have the same amplitude or are not ninety degrees out of phase, 
	though their phase offset and their amplitude ratio are constant.
This kind of polarization is called elliptical polarization because the electric vector traces out an ellipse in the plane (the polarization ellipse). 
%This is shown in the above figure on the right.

The parameters that describe the polarization ellipse are:
\begin{equation}
\tan (\theta ) = \frac{E_y^0}{E_x^0}
\end{equation}
with
	$0 \leq \theta \leq \pi/2$. 
The orientation angle, which is the angle between the major semi-axis of the ellipse and the x-axis, is given by:
\begin{equation}
\tan (\Psi) = \tan (2 \theta) \cos(\Delta \phi)
\end{equation}
with
	$0 \leq \Psi < \pi$. 
The ellipticity, $\epsilon$ is the major-to-minor-axis ratio.
The ellipticity angle which is defined as $\chi = \arctan (1/\epsilon)$ can be calculated as:
\begin{equation}
\sin (2 \chi) = \sin(2 \theta) \sin( \Delta \phi)
\end{equation}
with
	$0 \leq \chi \leq \pi/2$, and $0 \leq \epsilon \leq \infty$

Ignoring the scale of the electrical fields, the elementary wave polarization can be completely described by two parameters, i.e. $E_y^0/E_x^0$ and $\Delta \phi = \phi_y^0 - \phi_x^0$. 
Similarly, ignoring the scale of the ellipse, the polarization ellipse can be completely described by two angles $\Psi$ and $\chi$.

%The choice of basis vector used to represent horizontal and vertical directions are evidently arbitrary.
%In general, any pair of orthogonal polarization states can be used.
%Formally orthogonal vector pair is defined as one having a zero inner product.
%Another common choice is left and right circular.
%In anyway, implicit in the representations above is the orientation of the coordination frame.
%Of the two angles frequently used in polarization ellipse analysis, one $\Psi$ is dependent on the coordination frame and the other may not $\chi$

%To recap $\chi = \tan^{-1} (1/\epsilon)$, with $\epsilon$ being the ellipse's ellipticity.
%In polarization ellipse, $\epsilon=1$ indicates circular polarization, while $\epsilon=0$ or $\epsilon=\infty$ corresponds to linear polarizations.
%An alternative to ellipticity angle $\chi$ is the eccentricity angle.
%However, the use of ecccentricity angle has not been widely used, as it has no obvious geometrical interpretation in Poincare Spheres.

%The polarization of light can be measured with a polarimeter.

\subsubsection{Mathematical representations of wave polarizations}

\paragraph{ Complex polarization ratio }
This ratio is important in explaining polarization basis changes.

It is defined as:
\begin{equation}
\rho = \frac{E_y^0}{E_x^0} e^{i \Delta \phi} = \frac{\cos(2 \chi) + i \sin(2 \chi)}{1 - \cos(2\Psi) \cos(2\chi)}
\end{equation}

\paragraph{Jones Vector}

%Let the total wave power:
%\begin{equation}
%|E|^2 = (E_x^0)^2  + (E_y^0)^2
%\end{equation}
%and phase 
%\begin{equation}
%\theta = \tan^{-1} \left( \frac{E_y^0}{E_x^0} \right)
%\end{equation}

The Jones vector is defined as 
\begin{equation}
J = 
\left(
\begin{array}{c}
	\xi_x(t) \\
	\xi_y(t) 
\end{array}
\right)
= \xi_x(t)
\left(
\begin{array}{c}
	1 \\
	\rho
\end{array}
\right)
\end{equation}

Thus the normalized Jones vector can be given as:
\begin{equation}
J = 
\left(
\begin{array}{c}
	\cos(\theta) \cdot e^{i \phi_x^0} \\
	\sin(\theta) \cdot e^{i \phi_y^0}
\end{array}
\right)
=
\left(
\begin{array}{c}
	1\\
	\tan(\theta) \cdot e^{i \Delta \phi}
\end{array}
\right)
=
\left(
\begin{array}{c}
	\cos(\Psi) \cos(\chi) - i \sin(\Psi) \sin(\chi) \\
	\sin(\Psi) \cos(\chi) + i \cos(\Psi) \sin(\chi)
\end{array}
\right)
\end{equation}

The two parameters in the Jones vector can completely describe the polarization state (ellipse) of any elementary wave.
This same Jones vector finds extensive use in quantum machenics, and is interpreted as quantum state vector.
Its bra-ket notation formula, which is normally used in quantum computation contexts is:
\begin{equation}
J = | \Psi \rangle 
= 
\left(
\begin{array}{c}
	\cos(\theta) \cdot e^{i \phi_x^0} \\
	\sin(\theta) \cdot e^{i \phi_y^0}
\end{array}
\right)
\end{equation}

By definitions, the horizontal linear polarized wave is defined as: $J_h = \left(
\begin{array}{c}
 1 \\
 0
\end{array}
\right)$, similarly the vertical linear polarized wave: $J_v = \left(
\begin{array}{c}
 0 \\
 1
\end{array}
\right)$, the $\pm \pi/4$ linear polarized waves $J_{\pm} = \pm \frac{1}{\sqrt{2}} \left(
\begin{array}{c}
 1 \\
 1
\end{array}
\right)$, the left circular polarized wave: $J_l = |L\rangle = \frac{1}{\sqrt{2}} \left(
\begin{array}{c}
 1 \\
 -i
\end{array}
\right)$ and the right circular polarized wave: $J_r = |R\rangle = \frac{1}{\sqrt{2}} \left(
\begin{array}{c}
 1 \\
 i
\end{array}
\right)$

%Then any arbitrary state, written in R-L basis is given by:
%\begin{equation}
%| \Psi \rangle 
%= 
%a_R e^{i \phi_x - i \theta} | R \rangle 
%+ 
%a_L e^{i \phi_x + i \theta} | L \rangle 
%\end{equation}

\paragraph{ Stokes Vector }

@@

\subsection{Different Modes of Polarimetric SAR}

@@

\subsubsection{ Partial Polarimetry and the traditional SAR}

For the tasks of target detection and classification, we believe that partial dual polarization provides a unique offer.
Compared with single-channel, dual channel provide extra polarimetric information. 
Compared with full polarimetric data, the better spatial resolution could makes up for the loss of reconstructed polarimetric information, especially for the purpose of target detection and classification.

In fact, this project is intended to provide help for a new project at EADS Innovation Works Singapore to investigate the use of polarimertic data in ship monitoring over the oceans.
It is hypothesized that using polarimetric information would improve the possibilities of detecting ships travelling in the open seas.
Evidently partial polarimetric modes are good candidates for better detection rates.

The scientific application of POLSAR include: 
	oil slick detection \cite{Nunziata_2009_IGARSS, Gambardella_2007_ESASP}, 
	ship monitoring \cite{Kozai_2004_OCEANS}, 
	sea ice monitoring \cite{Wakabayashi_2004_TGRS, Nakamura_2005_IGARSS}, 
	volcano lava-flow monitoring \cite{Dierking_1998_IGARSS}, 
	agricultural crop monitoring \cite{Bouvet_2009_TGRS, Hajnsek_2008_IGARSS}, 
	planetary exploration \cite{Rosenbush_2009_JQuantSpecRadTransfer} 
	and other surface parameter estimations \cite{TruongLoi_2009_TGRS, Touzi_2009_TGRS, Hajnsek_Thesis_2001}. 
The remaining text of this chapter briefly look at recent literature in the context of ship detection and monitoring.
The topic of ship detection and monitoring has already been tackled quite extensively for single channel SAR data \cite{Tello_2005_IGARSS, Tello_2006_ISPRSJPhotogrammetryRemoteSensing}.
In fact, a number of countries have reported to implement such systems, and some uses them in everyday operations.
The countries includes: USA \cite{Montgomery_1998_IGARSS}, Canada \cite{Vachon_2000_JHUApplPhysics}, France \cite{Losekoot_2005_IGARSS}, Norway \cite{weydahl_2007_RemoteSenseGISGeology} and Spain \cite{Margarit_2009_RemoteSensing}.
The common scenarios in use are normally: 
	a designated area is continuously monitored through the use of some automatic ship detection algorithm.
If an unknown ship is detected, some manual intervention is to be applied, normally in the form of a navy ship being sent to further investigate.
The reports however did not elaborate about the accuracy of the detection algorithms.

A recent trend that is reported in the literature is the use of AIS (Automatic Identification System).
In 2007, Vachon \cite{Vachon_2007_IGARSS} reported his study in Canada, which aimed to match ships detectable in SAR images with AIS reported locations.
In 2009, Grasso \cite{Grasso_2009_ISDA} evaluated the performance of ship detection algorithms using AIS as ground truth.
In our planned approach, we are prepared to study the performance improvement through using partial polarimetry data in comparison to normal single-channel SAR approaches.
Ground truth for such study will also be using synchronized AIS data.

\section{The Stochastics and Multivariate Nature of (POL)SAR data}

%\subsection{The Stochastics Nature of SAR and POLSAR}
\subsection{The Stochastic nature of SAR  and SAR speckle filtering }

This section starts with a description of the stochastic nature of SAR data. 
SAR speckle filtering is then reviewed within a statistical estimation theory framework. 
Various different approaches to SAR speckle filtering are reviewed, 
	ranging from Generalized Least Squared Error algorithms
	%ranging from a discrete cartoon like algorithm 
	to the more convoluted Maximum A Posteriori (MAP) approaches.
	%to recent more continous multi-resolution wavelet or partial differental equations based algorithms.

@@

\subsection{The Multivariate Nature and The Observables in POLSAR data}

\section{Current methods in SAR and POLSAR speckle filtering}

\subsection{Current Methods for SAR speckle filtering}
(2.2.1)

@@

\section{Existing Measures of Distance for (POL)SAR data and their applications}

@@

\section{The need for Scalar Consistent Measures of Distance}
	We will show our originality and significance here!

>>>>

\section{The Stochastics and Multivariate Nature of POLSAR data}

\subsection{The Stochastic nature of SAR data}

\subsubsection{ Traditional Single-Channel SAR and the Speckle phenomena}

Wikipedia to be edited strongly.
Speckle noise is a granular noise that inherently exists in and degrades the quality of the active radar and synthetic aperture radar (SAR) images.
Speckle noise in conventional radar results from random fluctuations in the return signal from an object that is no bigger than a single image-processing element. It increases the mean grey level of a local area.[1]
Speckle noise in SAR is generally more serious, causing difficulties for image interpretation.[1][2] It is caused by coherent processing of backscattered signals from multiple distributed targets. In SAR oceanography, for example, speckle noise is caused by signals from elementary scatterers, the gravity-capillary ripples, and manifests as a pedestal image, beneath the image of the sea waves.[3][4]

\subsubsection{The Stochastic nature of SAR  and SAR speckle filtering }

The stochastic nature SAR arises from the speckle phenomena.
As SAR speckle seriously hinders the interpretation and understanding of SAR images,
  its effect are routinely reduced through a speckle filtering process.
SAR speckle filtering is the process to suppress the effect of speckle noise SAR images for easier intepretation.

\subsection{The Multivariate Nature of POLSAR data}

\subsubsection{ Partial Polarimetry and the Stokes Vector}

For the tasks of target detection and classification, we believe that partial dual polarization provides a unique offer.
Compared with single-channel, dual channel provide extra polarimetric information. 
Compared with full polarimetric data, the better spatial resolution could makes up for the loss of reconstructed polarimetric information, especially for the purpose of target detection and classification.
\paragraph{ Stokes Vector }

@@

%In fact, this project is intended to provide help for a new project at EADS Innovation Works Singapore to investigate the use of polarimertic data in ship monitoring over the oceans.
%It is hypothesized that using polarimetric information would improve the possibilities of detecting ships travelling in the open seas.
%Evidently partial polarimetric modes are good candidates for better detection rates.
%
%The scientific application of POLSAR include: 
%	oil slick detection \cite{Nunziata_2009_IGARSS, Gambardella_2007_ESASP}, 
%	ship monitoring \cite{Kozai_2004_OCEANS}, 
%	sea ice monitoring \cite{Wakabayashi_2004_TGRS, Nakamura_2005_IGARSS}, 
%	volcano lava-flow monitoring \cite{Dierking_1998_IGARSS}, 
%	agricultural crop monitoring \cite{Bouvet_2009_TGRS, Hajnsek_2008_IGARSS}, 
%	planetary exploration \cite{Rosenbush_2009_JQuantSpecRadTransfer} 
%	and other surface parameter estimations \cite{TruongLoi_2009_TGRS, Touzi_2009_TGRS, Hajnsek_Thesis_2001}. 
%The remaining text of this chapter briefly look at recent literature in the context of ship detection and monitoring.
%The topic of ship detection and monitoring has already been tackled quite extensively for single channel SAR data \cite{Tello_2005_IGARSS, Tello_2006_ISPRSJPhotogrammetryRemoteSensing}.
%In fact, a number of countries have reported to implement such systems, and some uses them in everyday operations.
%The countries includes: USA \cite{Montgomery_1998_IGARSS}, Canada \cite{Vachon_2000_JHUApplPhysics}, France \cite{Losekoot_2005_IGARSS}, Norway \cite{weydahl_2007_RemoteSenseGISGeology} and Spain \cite{Margarit_2009_RemoteSensing}.
%The common scenarios in use are normally: 
%	a designated area is continuously monitored through the use of some automatic ship detection algorithm.
%If an unknown ship is detected, some manual intervention is to be applied, normally in the form of a navy ship being sent to further investigate.
%The reports however did not elaborate about the accuracy of the detection algorithms.
%
%A recent trend that is reported in the literature is the use of AIS (Automatic Identification System).
%In 2007, Vachon \cite{Vachon_2007_IGARSS} reported his study in Canada, which aimed to match ships detectable in SAR images with AIS reported locations.
%In 2009, Grasso \cite{Grasso_2009_ISDA} evaluated the performance of ship detection algorithms using AIS as ground truth.
%In our planned approach, we are prepared to study the performance improvement through using partial polarimetry data in comparison to normal single-channel SAR approaches.
%Ground truth for such study will also be using synchronized AIS data.

@@

%\subsection{The Stochastics Nature of SAR and POLSAR}
\section{Current methods in SAR and POLSAR Data Processing}

\subsection{Current Methods for SAR speckle filtering}
This section starts with a description of the stochastic nature of SAR data. 
SAR speckle filtering is then reviewed within a statistical estimation theory framework. 
Various different approaches to SAR speckle filtering are reviewed, 
	ranging from Generalized Least Squared Error algorithms
	%ranging from a discrete cartoon like algorithm 
	to the more convoluted Maximum A Posteriori (MAP) approaches.
	%to recent more continous multi-resolution wavelet or partial differental equations based algorithms.

@@

@@

\subsection{Existing Dis-similarity Measures for (POL)SAR data}

@@

\subsection{The need for Scalar Consistent Measures of Distance}
	We will show our originality and significance here!
